---
title: "MKTG 366 Project - LambdaMART"
output: word_document
toc: yes
---


```{r, echo=FALSE, results=FALSE, warning=FALSE}
setwd("~/Google Drive/School/MKTG 366 Project/expedia-hotel-search/")
library(data.table)
library(caret)
library(dplyr)
library(gbm)
```


## Data preprocessing

```{r, results=FALSE}
# load data
exped = data.table::fread("train.csv")
# us_data = data.table::fread("train.csv")
test_data = fread("../test.csv")
```

In the following we construct 3 variables: srch\_month, travel\_month, and int\_travel.
```{r}
exped$srch_month = factor(month(exped$date_time) )
exped$travel_month = factor(month(as.Date(exped$date_time) + exped$srch_booking_window))
exped$intl_travel = ifelse(exped$visitor_location_country_id != 219, 1, 0)
```


The following shows the glimps of the data and helps us decide what pre-processing operations to take.

```{r, warning=FALSE}
str(data_us)
```


We remove the following variables:

- visitor\_hist_staring, visitor\_hist\_adr\_usd, srch\_query\_affinity\_score, gross\_bookings\_usd, since for the majority of observations they are NULL
- all comp\_ variables, since they are mostly NULL and it has been observed that these variables do not have significant impact on booking
- prop\_country\_id, since it's 219 for all examples
- srch\_destination\_id, since it has too many levels for gbm to handle (max 1024 levels)

```{r, warning=FALSE}
data_us = exped %>% select(click_bool, random_bool, booking_bool, prop_brand_bool,
                             srch_id, site_id, visitor_location_country_id, #prop_id,
                             prop_starrating, prop_review_score, prop_location_score1, prop_location_score2,
                             prop_log_historical_price, price_usd, #position,
                             srch_length_of_stay, srch_booking_window, orig_destination_distance,
                             srch_adults_count, srch_children_count, srch_room_count,
                             srch_month, travel_month, intl_travel)
```

Of the variables selected, we convert:

- all \_bool variables and all \_id (except srch\_id) variables to "factor"
- prop\_review_score, prop\_location\_score2, orig\_destination_distance from "chr" to "numeric"

```{r, warning=FALSE}
# convert variables that were coerced to "character" to type "numeric"
numeric_names = c("prop_review_score", "prop_location_score2", "orig_destination_distance")
setDT(data_us)[, (numeric_names) := lapply(.SD, as.numeric), .SDcols = numeric_names]

# convert these variables to type "factor"
bool_names = c("click_bool", "prop_brand_bool", "booking_bool", "random_bool")
id_names = c("site_id", "visitor_location_country_id")
setDT(data_us)[, (bool_names) := lapply(.SD, as.factor), .SDcols = bool_names]
setDT(data_us)[, (id_names) := lapply(.SD, as.factor), .SDcols = id_names]
```

We create the response variable as a composition of variables **booking\_bool** and **click\_bool**. A booking is scored highest at 5, a click is scored at 1, and no action is scored 0. This captures the value of the corresponding action.

```{r, warning=FALSE}
data_us$response = ifelse(data_us$booking_bool == 1, 5, ifelse(data_us$click_bool == 1, 1, 0))
data_us = data_us %>% select(c(-booking_bool, -click_bool))
```


Proportion of observations that are NA, for each variable:
```{r, warning=FALSE}
na.proportion = function(column){
  return(sum(is.na(column)) / length(column))
}

t = lapply(data_us, na.proportion)
t
```
So we are fine.


Similar processing is performed on the test set


```{r}
test_data$srch_month = factor(month(test_data$date_time) )
test_data$travel_month = factor(month(as.Date(test_data$date_time) + test_data$srch_booking_window))
test_data$intl_travel = ifelse(test_data$visitor_location_country_id != 219, 1, 0)
```

```{r}
test_clean = test_data %>% select(prop_brand_bool,
                                   srch_id, site_id, visitor_location_country_id, prop_id,
                             prop_starrating, prop_review_score, prop_location_score1, prop_location_score2,
                             prop_log_historical_price, price_usd,
                             srch_length_of_stay, srch_booking_window, orig_destination_distance,
                             srch_adults_count, srch_children_count, srch_room_count,
                             srch_month, travel_month, intl_travel)
```

```{r}
numeric_names = c("prop_review_score", "prop_location_score2", "orig_destination_distance")
setDT(test_clean)[, (numeric_names) := lapply(.SD, as.numeric), .SDcols = numeric_names]

# convert these variables to type "factor"
id_names = c("site_id", "visitor_location_country_id")
setDT(test_clean)[, (id_names) := lapply(.SD, as.factor), .SDcols = id_names]
```


## Modeling

In the following we run a tree ensemble modeling using LambdaMART algorithm to learn to rank hotel results in response to a query.

We train on the full dataset, but only on results that were sorted by relevance by Expedia (i.e. ```random\_bool == 0```).


```{r}
train_not_random = data_us %>% filter(random_bool == 0) %>% select(-random_bool)

lambdamart_not_random1000 = gbm(response ~ .,
              data = train_not_random,
              distribution = list(name = 'pairwise', metric = "ndcg", group = 'srch_id'),
              n.trees = 1000,
              interaction.depth = 4,
              shrinkage = 0.005,
              train.fraction = 0.5,
              verbose = TRUE)

summary(
  lambdamart_not_random,
  n.trees = gbm.perf(lambdamart_not_random),
  main = 'pairwise (ndcg)'
  )

```

Variable importance for the most part match the exploratory conclusions we’ve reached earlier. For instance, position of a hotel in the ranking  is the most influential on whether that hotel is clicked or booked.


## Making prediction

Step 1: use ```type = 'response'``` to get predicted probabilities of the results.
```{r, warning=FALSE}
predictions = predict(lambdamart_not_random,
                      test_clean,
                      type = 'response',
                      n.trees = gbm.perf(lambdamart_not_random))
```

Step 2: Combine search id, predicted probabilities, and property id in one data frame
```{r}
preds = data.frame(cbind(test_clean$srch_id, predictions, test_clean$prop_id))
colnames(preds) = c("srch_id", "predicted_prob", "prop_id")
```

Step3: Order the results first by srch_id in ascending order, then by predicted probability in descending order (for each search id)

Use DATA.TABLE's setorder takes less than 2 seconds. Other operations using dplyr or core R will take several hours.
```{r}
ranked_results_2 = setorder(setDT(preds), srch_id, -predicted_prob)
```



## Writing predictions to file

```{r}
output_to_save = ranked_results_2 %>% select(srch_id, prop_id)
colnames(output_to_save) = c("SearchId", "PropertyId")
write.csv(output_to_save, "submission.csv", sep = ",", row.names = FALSE)
```

In terminal, type the following to correct the header (there should not be quotes around the column names):

```sed -i -e "1d" submission.csv```

```sed -i 1i"SearchId,PropertyId" submission.csv```
